Weighted Model Counting Without Parameter Variables

### 2021-02-26 doc/paper3/paper.tex
Writing and editing

% TODO: Now say that our work is similar: it introduces a new representation
% that is more appropriate for algorithms such as DPMC. Stress that all initial
% pseudo-Boolean functions have two values.

@@

\begin{remark}
  Encodings such as \texttt{cd05} and \texttt{cd06} are \emph{not} WMC
  encodings. Instead, they encode Bayesian network inference into instances of
  the \emph{minimum-cardinality} WMC problem, where the answer is defined to be
  $\sum_{Y \models \phi\text{, }|Y| = k} \prod_{Y \models l}
  w(l)$, where $k = \min_{Y \models \phi\text{, }Y \ne \emptyset}
  |Y|$, if $k$ exists, otherwise the answer is zero. This additional
  condition on model cardinality becomes necessary because these encodings
  eliminate clauses of the form $p \Rightarrow i$, where $p \in X_P$ is a
  parameter variable, and $i \in X_I$ is an indicator variable. Nonetheless, our
  transformation algorithm still works on such encodings, although the
  experimental results are discouraging because they use approximately twice as
  many indicator variables. For instance, each binary variable of a Bayesian
  network is encoded using two indicator variables while one would suffice.
\end{remark}

\begin{definition}[PBP Instance] \label{def:new_wmc}
  A \emph{pseudo-Boolean projection} (PBP) instance is a tuple $(F, X, \omega)$,
  where $X$ is the set of variables, $F$ is a set of pseudo-Boolean functions
  $2^X \to \mathbb{R}$, and $\omega \in \mathbb{R}$ is the scaling factor. The
  \emph{answer} of the instance is $\omega \cdot \left(\exists_X\prod_{f \in
      F}f\right)(\emptyset)$.
\end{definition}

NOTE: The constant is inspired by the \texttt{bklm16}
\cite{DBLP:conf/ecai/BartKLM16} encoding.

@@

\begin{itemize}
\item \texttt{d02} \cite{DBLP:conf/kr/Darwiche02}
\item \texttt{sbk05} \cite{DBLP:conf/aaai/SangBK05}
\item \texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05}
\item \texttt{cd06} \cite{DBLP:conf/sat/ChaviraD06}
\item \texttt{bklm16} \cite{DBLP:conf/ecai/BartKLM16}
\end{itemize}

@@

\begin{lemma} \label{lemma:sum}
  For any pseudo-Boolean function $f\colon 2^X \to \mathbb{R}$,
  \[
    (\exists_X f)(\emptyset) = \sum_{Y \subseteq X} f(Y).
  \]
\end{lemma}
\begin{proof}
  For base case, let $X = \{x\}$. Then
  \[
    (\exists_xf)(\emptyset) = (f|_{x=1} + f|_{x=0})(\emptyset) =
    f|_{x=1}(\emptyset) + f|_{x=0}(\emptyset) = \sum_{Y \subseteq \{x\}} f(Y).
  \]
  This easily extends to $|X| > 1$ by the definition of projection on sets of
  variables.
\end{proof}

% NOTE: mention that this is similar (but not equal) to the DPMC/ADDMC result
\begin{proposition} \label{prop:equivalence}
  Let $(\phi, X_I, X_P, w)$ be a WMC instance. Then
  \begin{equation}
  \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
        x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
    1\right) \label{eq:new_wmc}
  \end{equation}
  is a PBP instance with the same answer.
\end{proposition}
\begin{proof}
  The answer of \cref{eq:new_wmc} is
  \begin{align*}
    \left(\exists_{X_I \cup X_P} \left(\prod_{c \in \phi} [c]_0^1\right) \prod_{x \in X_I \cup X_P} [x]_{w(\neg x)}^{w(x)} \right)(\emptyset) &= \sum_{Y \subseteq X_I \cup X_P} \left(\left(\prod_{c \in \phi} [c]_0^1 \right) \prod_{x \in X_I \cup X_P} [x]_{w(\neg x)}^{w(x)}\right)(Y) \\
&= \sum_{Y \subseteq X_I \cup X_P} \underbrace{\left(\prod_{c \in \phi} [c]_0^1\right)}_f(Y) \underbrace{\left(\prod_{x \in X_I \cup X_P} [x]_{w(\neg x)}^{w(x)}\right)}_g(Y)
  \end{align*}
  by \cref{lemma:sum}. Note that
  \[
    f(Y) =
    \begin{cases}
      1 & \text{if } Y \models \phi, \\
      0 & \text{otherwise},
    \end{cases}
    \quad
    \text{and}
    \quad
    g(Y) = \prod_{Y \models l} w(l),
  \]
  which means that
  \[
    \sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y) = \sum_{Y \models \phi} \prod_{Y
      \models l} w(l)
  \]
  as required.
\end{proof}

@@

  Let us first consider how the unfinished WMC instance $(F, X_I, \omega)$ after
  the loop on \crefrange{line:foreach1start}{line:foreach1end} differs from
  \eqref{eq:new_wmc2}. Note that \cref{alg:transformation} leaves each $c \in
  \phi_*$ unchanged, i.e., adds $[c]_0^1$ to $F$. We can then fix an arbitrary
  $p \in X_P$ and let $F_p$ be the set of functions added to $F$ as a
  replacement of $\phi_p$. It is sufficient to show that
  \begin{equation} \label{eq:to_show}
    \omega \prod_{f \in F_p} f = \exists_p [p]_{w(\neg p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1.
  \end{equation}
  Note that under \cref{condition:d02},
  \[
    \bigwedge_{c \in \phi_p} c \equiv p \Leftrightarrow \bigwedge_{i=1}^n l_i
  \]
  for some family of indicator variable literals $(l_i)_{i=1}^n$. Thus,
  \[
    \exists_p [p]_{w(\neg p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 = \exists_p
    [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1.
  \]
  If $w(p) = 1$, then
  \begin{equation} \label{eq:bigsums}
    \exists_p [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i
    \right]_0^1 = \exists_p \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i
    \right]_0^1 = \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i
    \right]_0^1\right|_{p=1} + \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n
      l_i \right]_0^1\right|_{p=0}.
  \end{equation}
  Since for any input, $\bigwedge_{i=1}^n l_i$ is either true or false, exactly
  one of the two summands in \cref{eq:bigsums} will be equal to one, and the
  other will be equal to zero, and so
  \[
    \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i
      \right]_0^1\right|_{p=1} + \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n
        l_i \right]_0^1\right|_{p=0} = 1,
  \]
  where $1$ is a pseudo-Boolean function that always returns one. On the other
  side of \cref{eq:to_show}, since $F_p = \emptyset$, and $\omega$ is unchanged,
  we get $\omega\prod_{f \in F_p} f = 1$, and so \cref{eq:to_show} is satisfied
  under \cref{condition:d02} when $w(p) = 1$.

  If $w(p) \ne 1$, then
  \[
    F_p = \left\{ \left[ \bigwedge_{i = 1}^n l_i \right]_1^{w(p)} \right\},
  \]
  and $\omega = 1$, and so we want to show that
  \[
    \left[ \bigwedge_{i = 1}^n l_i \right]_1^{w(p)} = \exists_p [p]_1^{w(p)}
    \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1,
  \]
  and indeed
  \begin{align*}
    \exists_p [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1 &= \left.\left( [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1 \right)\right|_{p=1} + \left.\left([p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1 \right)\right|_{p=0} \\
                                                                                      &= w(p) \cdot \left[ \bigwedge_{i=1}^n l_i \right]_0^1 + \left[ \bigwedge_{i=1}^n l_i \right]_1^0 = \left[ \bigwedge_{i=1}^n l_i \right]_1^{w(p)}.
  \end{align*}
  This finishes the proof of the correctness of the first `foreach' loop under
  \cref{condition:d02}.

@@

  If $w(p) \ne 1$, and $\{ p \} \in \phi_p$, then, by
  \cref{cond:just_parameter}, $\phi_p = \{ \{ p \} \}$, and
  \cref{alg:transformation} produces $F_p = \emptyset$ and $\omega = w(p)$, and
  so
  \[
    \exists_p [p]_{w(\neg p)}^{w(p)} [p]_0^1 = \exists_p [p]^{w(p)}_0 = w(p) =
    \omega \prod_{f \in F_p} f.
  \]
  The only remaining case is when $w(p) \ne 1$ and $\{ p \} \not \in \phi_p$.
  Then $\omega = 1$, and
  \[
    F_p = \left\{ \left[\bigwedge_{l \in c \setminus \{ p \}} \neg
        l\right]_1^{w(p)} \;\middle|\; c \in \phi_p \right\},
  \]
  so need to show that
  \[
    \prod_{c \in \phi_p} \left[\bigwedge_{l \in c \setminus \{ p \}} \neg
      l\right]_1^{w(p)} = \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p}
    [c]_0^1.
  \]
  We can rearrange the right-hand side as
  \begin{align*}
    \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 &= \exists_p [p]_{1-w(p)}^{w(p)} \left[\bigwedge_{c \in \phi_p} c\right]_0^1 = \exists_p [p]_{1-w(p)}^{w(p)} \left[ p \lor \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
                                                               &= w(p) + (1-w(p)) \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 = w(p) + \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^{1-w(p)} \\
                                                               &= \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_{w(p)}^1 = \left[ \neg \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \neg(c \setminus \{ p \}) \right]_1^{w(p)} \\
                                                               &= \left[ \bigvee_{c \in \phi_p} \neg \bigvee_{l \in c \setminus \{ p \}} l \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l \right]_1^{w(p)}.
  \end{align*}
   By \cref{cond:disjoint}, $\bigwedge_{l \in c \setminus \{ p \}} \neg l$ can
   be true for at most one $c \in \phi_p$, and so
   \[
     \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l
     \right]_1^{w(p)} = \prod_{c \in \phi_p} \left[ \bigwedge_{l \in c \setminus
         \{ p \}} \neg l \right]_1^{w(p)}
   \]
   which is exactly what we needed to show. This ends the proof that the
   first loop of \cref{alg:transformation} preserves the answer under both
   \cref{condition:d02} and \cref{condition:bklm16}. Finally, the loop on
   \crefrange{line:foreach2start}{line:foreach2end} of \cref{alg:transformation}
   replaces $[v]_1^p[\neg v]_1^q$ with $[v]_q^p$ (for some $v \in X_I$ and $p, q
   \in \mathbb{R}$), but, of course,
   \[
     [v]_1^p[\neg v]_1^q = [v]_1^p[v]_q^1 = [v]_q^p,
   \]
   i.e., the answer is unchanged.
\end{proof}

@@

\paragraph{Notes.}
\begin{itemize}
\item The experiments were run on a computing cluster with Intel Xeon E5-2630,
  Intel Xeon E7-4820, and Intel Xeon Gold 6138 processors with a
  \SI{1000}{\second} timeout separately on both encoding and inference, and a
  \SI{32}{\gibi\byte} memory limit.\footnote{Each instance was run on the same
    processor across all algorithms and encodings.}
\item Along with
  \textsf{DPMC}\footnote{\url{https://github.com/vardigroup/DPMC}}, in our
  experiments we also include WMC algorithms used in the papers that introduce
  each encoding: \textsf{Ace}
  for \texttt{cd05}, \texttt{cd06}, and \texttt{d02};
  \textsf{Cachet}\footnote{\url{https://cs.rochester.edu/u/kautz/Cachet/}}
  \cite{DBLP:conf/sat/SangBBKP04} for \texttt{sbk05}; and
  \textsf{c2d}\footnote{\url{http://reasoning.cs.ucla.edu/c2d/}}
  \cite{DBLP:conf/ecai/Darwiche04} with
  \textsf{query-dnnf}\footnote{\url{http://www.cril.univ-artois.fr/kc/d-DNNF-reasoner.html}}
  for \texttt{bklm16}.
\item \textsf{DPMC} was run in \emph{greedy mode}. This mode (which was not part
  of the original paper \cite{DBLP:conf/cp/DudekPV20}) optimises the order in
  which pseudo-Boolean functions are multiplied by prioritising functions with
  small representations.
\item Indeed, the numbers of variables in \texttt{bklm16++} and \texttt{d02++}
  encodings are equal whenever all variables in a Bayesian network have two
  values (which applies to most of the benchmark instances). Nonetheless,
  \texttt{bklm16++} performs much faster than \texttt{d02++}.
\item \textsf{DPMC} is run with tree decomposition based planning and algebraic
  decision diagram based execution---the best-performing combination in the
  original set of experiments \cite{DBLP:conf/cp/DudekPV20}. We use a single
  iteration of \textsf{htd} \cite{DBLP:conf/cpaior/AbseherMW17} to generate
  approximately optimal tree decompositions---we found that this configuration
  is efficient enough to handle huge instances, and yet the width of the
  returned decomposition is unlikely to differ from optimal by more than one or
  two.
\item We also record numbers of variables in each encoding before and after our
  additional transformation.
\item Experimental research questions:
  \begin{itemize}
  \item Can parameter variable elimination improve inference speed?
  \item How does DPMC combined with encodings without (and with) parameter
    variables compare with other WMC algorithms and other encodings?
  \item Which instances is our approach particularly successful on (compared to
    other algorithms and encodings and to the same encoding before our
    transformation)?
  \item What proportion of variables is typically removed?
  \item Do some encodings benefit from this transformation more than others?
  \end{itemize}
\item How many times am I faster than old cd06 and new bklm16? Perhaps add other
  numbers as well.
\end{itemize}
% TODO: define binary = 2 values per variable (or states per node?)
% reuse this term where I'm referring to two values

>>>>

Our work is similar in that it introduces a new representation for computational
problems that is based on pseudo-Boolean functions---pseudo-Boolean projection
(PBP). We formally show that every WMC problem instance has a corresponding PBP
instance. We argue that this new format is more appropriate for WMC solvers
whose execution is build on pseudo-Boolean function manipulation, i.e.,
\textsf{ADDMC} \cite{DBLP:conf/aaai/DudekPV20} and \textsf{DPMC}
\cite{DBLP:conf/cp/DudekPV20}. Finally, we show how most methods of encoding
probabilistic inference on Bayesian networks to WMC can be adapted to the PBP
format. This transformation significantly improves inference speed,
outperforming all other considered algorithm-encoding combinations.

%allowing two encodings that were previously incompatible with
%\textsf{DPMC} to be compatible.

% TODO: we also show correctness
% TODO: and new notation!

@@

\begin{definition}[PBP Instance] \label{def:new_wmc}
  A PBP instance is a tuple $(F, X, \omega)$, where $X$ is the set of variables,
  $F$ is a set of pseudo-Boolean functions $2^X \to \mathbb{R}$, and $\omega \in
  \mathbb{R}$ is the scaling factor.\footnote{We will further restrict $F$ to
    only contain functions that have exactly two values. While this restriction
    is not necessary from a computational point of view, such functions can be
    conveniently represented both in text files and using the syntax proposed in
    this paper.} The \emph{answer} of the instance is $\omega \cdot
  \left(\exists_X\prod_{f \in F}f\right)(\emptyset)$.
\end{definition}

NOTE: Adding scaling factor $\omega$ to the definition allows us to remove
clauses that consist entirely of a single parameter variable. The idea of
extracting some of the structure of the WMC instance into an external
multiplicative factor originates in one of the Bayesian network encodings
\cite{DBLP:conf/ecai/BartKLM16}.

@@

\paragraph{TODO}
\begin{itemize}
\item Define the terminology surrounding BNs: variables, values, directed
  probabilistic graphical model, DAG, binary = two values per variable.
\item Reuse this term where I'm referring to two values.
\item Give a brief description of each encoding.
\item Talk about the parameter/indicator variable distinction.
\end{itemize}

WMC is a well-established technique for Bayesian network inference, particularly
effective on networks where most variables have only a few possible values
\cite{DBLP:conf/kr/Darwiche02}. Many ways of encoding a Bayesian network into a
WMC instance have been proposed. We will refer to them based on initials of the
authors and the year of publication: \texttt{d02}
\cite{DBLP:conf/kr/Darwiche02}, \texttt{sbk05} \cite{DBLP:conf/aaai/SangBK05},
\texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05}, \texttt{cd06}
\cite{DBLP:conf/sat/ChaviraD06}, \texttt{bklm16} \cite{DBLP:conf/ecai/BartKLM16}.

\begin{remark}
  Strictly speaking, encodings such as \texttt{cd05} and \texttt{cd06} are
  \emph{not} WMC encodings. Instead, they encode Bayesian network inference into
  instances of the \emph{minimum-cardinality} WMC problem, where the answer is
  defined to be $\sum_{Y \models \phi\text{, }|Y| = k} \prod_{Y \models l}
  w(l)$, where $k = \min_{Y \models \phi\text{, }Y \ne \emptyset} |Y|$, if $k$
  exists, otherwise the answer is zero. This additional condition on model
  cardinality becomes necessary because these encodings eliminate clauses of the
  form $p \Rightarrow i$, where $p \in X_P$ is a parameter variable, and $i \in
  X_I$ is an indicator variable. Nonetheless, our transformation algorithm still
  works on such encodings, although the experimental results are discouraging
  because they use approximately twice as many indicator variables. For
  instance, each binary variable of a Bayesian network is encoded using two
  indicator variables while one would suffice.
\end{remark}

@@

\begin{lemma} \label{lemma:sum}
  For any pseudo-Boolean function $f\colon 2^X \to \mathbb{R}$,
  \[
    (\exists_X f)(\emptyset) = \sum_{Y \subseteq X} f(Y).
  \]
\end{lemma}
\begin{proof}
  If $X = \{x\}$, then
  \[
    (\exists_xf)(\emptyset) = (f|_{x=1} + f|_{x=0})(\emptyset) =
    f|_{x=1}(\emptyset) + f|_{x=0}(\emptyset) = \sum_{Y \subseteq \{x\}} f(Y).
  \]
  This easily extends to $|X| > 1$ by the definition of projection on sets of
  variables.
\end{proof}

% NOTE: mention that this is similar (but not equal) to the DPMC/ADDMC result
\begin{proposition} \label{prop:equivalence}
  Let $(\phi, X_I, X_P, w)$ be a WMC instance. Then
  \begin{equation}
  \left(\left\{[c]_0^1 \;\middle|\; c \in \phi\right\} \cup \left\{[x]_{w(\neg
        x)}^{w(x)} \;\middle|\; x \in X_I \cup X_P\right\}, X_I \cup X_P,
    1\right) \label{eq:new_wmc}
  \end{equation}
  is a PBP instance with the same answer.
\end{proposition}
\begin{proof} % TODO: equation is not the right word
  Let $f = \prod_{c \in \phi} [c]_0^1$, and $g = \prod_{x \in X_I \cup X_P}
  [x]_{w(\neg x)}^{w(x)}$. Then the answer of \cref{eq:new_wmc} is
  \[
    (\exists_{X_I \cup X_P} fg)(\emptyset) = \sum_{Y \subseteq X_I \cup X_P}
    (fg)(Y) = \sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y)
  \]
  by \cref{lemma:sum}. Note that
  \[
    f(Y) =
    \begin{cases}
      1 & \text{if } Y \models \phi, \\
      0 & \text{otherwise},
    \end{cases}
    \quad
    \text{and}
    \quad
    g(Y) = \prod_{Y \models l} w(l),
  \]
  which means that
  \[
    \sum_{Y \subseteq X_I \cup X_P} f(Y)g(Y) = \sum_{Y \models \phi} \prod_{Y
      \models l} w(l)
  \]
  as required.
\end{proof}

@@

  Let us first consider how the unfinished WMC instance $(F, X_I, \omega)$ after
  the loop on \crefrange{line:foreach1start}{line:foreach1end} differs from
  \eqref{eq:new_wmc2}. Note that \cref{alg:transformation} leaves each $c \in
  \phi_*$ unchanged, i.e., adds $[c]_0^1$ to $F$. We can then fix an arbitrary
  $p \in X_P$ and let $F_p$ be the set of functions added to $F$ as a
  replacement of $\phi_p$. It is sufficient to show that
  \begin{equation} \label{eq:to_show}
    \omega \prod_{f \in F_p} f = \exists_p [p]_{w(\neg p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1.
  \end{equation}
  Note that under \cref{condition:d02},
  \[
    \bigwedge_{c \in \phi_p} c \equiv p \Leftrightarrow \bigwedge_{i=1}^n l_i
  \]
  for some family of indicator variable literals $(l_i)_{i=1}^n$. Thus,
  \[
    \exists_p [p]_{w(\neg p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 = \exists_p
    [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1.
  \]
  If $w(p) = 1$, then
  \begin{equation} \label{eq:bigsums}
    \exists_p [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1 = \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1\right|_{p=1} + \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1\right|_{p=0}.
  \end{equation}
  Since for any input, $\bigwedge_{i=1}^n l_i$ is either true or false, exactly
  one of the two summands in \cref{eq:bigsums} will be equal to one, and the
  other will be equal to zero, and so
  \[
    \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i
      \right]_0^1\right|_{p=1} + \left.\left[ p \Leftrightarrow \bigwedge_{i=1}^n
        l_i \right]_0^1\right|_{p=0} = 1,
  \]
  where $1$ is a pseudo-Boolean function that always returns one. On the other
  side of \cref{eq:to_show}, since $F_p = \emptyset$, and $\omega$ is unchanged,
  we get $\omega\prod_{f \in F_p} f = 1$, and so \cref{eq:to_show} is satisfied
  under \cref{condition:d02} when $w(p) = 1$.

  If $w(p) \ne 1$, then
  \[
    F_p = \left\{ \left[ \bigwedge_{i = 1}^n l_i \right]_1^{w(p)} \right\},
  \]
  and $\omega = 1$, and so we want to show that
  \[
    \left[ \bigwedge_{i = 1}^n l_i \right]_1^{w(p)} = \exists_p [p]_1^{w(p)}
    \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i \right]_0^1,
  \]
  and indeed
  \[
    \exists_p [p]_1^{w(p)} \left[ p \Leftrightarrow \bigwedge_{i=1}^n l_i
    \right]_0^1 = w(p) \cdot \left[ \bigwedge_{i=1}^n l_i \right]_0^1 + \left[
      \bigwedge_{i=1}^n l_i \right]_1^0 = \left[ \bigwedge_{i=1}^n l_i
    \right]_1^{w(p)}.
  \]
  This finishes the proof of the correctness of the first `foreach' loop under
  \cref{condition:d02}.

@@

  If $w(p) \ne 1$, and $\{ p \} \in \phi_p$, then, by
  \cref{cond:just_parameter}, $\phi_p = \{ \{ p \} \}$, and
  \cref{alg:transformation} produces $F_p = \emptyset$ and $\omega = w(p)$, and
  so
  \[
    \exists_p [p]_{w(\neg p)}^{w(p)} [p]_0^1 = \exists_p [p]^{w(p)}_0 = w(p) =
    \omega \prod_{f \in F_p} f.
  \]
  The only remaining case is when $w(p) \ne 1$ and $\{ p \} \not \in \phi_p$.
  Then $\omega = 1$, and
  \[
    F_p = \left\{ \left[\bigwedge_{l \in c \setminus \{ p \}} \neg
        l\right]_1^{w(p)} \;\middle|\; c \in \phi_p \right\},
  \]
  so need to show that
  \[
    \prod_{c \in \phi_p} \left[\bigwedge_{l \in c \setminus \{ p \}} \neg
      l\right]_1^{w(p)} = \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p}
    [c]_0^1.
  \]
  We can rearrange the right-hand side as % TODO: make it fit into margins
  \begin{align*}
    \exists_p [p]_{1-w(p)}^{w(p)} \prod_{c \in \phi_p} [c]_0^1 &= \exists_p [p]_{1-w(p)}^{w(p)} \left[\bigwedge_{c \in \phi_p} c\right]_0^1 = \exists_p [p]_{1-w(p)}^{w(p)} \left[ p \lor \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 \\
                                                               &= w(p) + (1-w(p)) \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^1 = w(p) + \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_0^{1-w(p)} \\
                                                               &= \left[ \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_{w(p)}^1 = \left[ \neg \bigwedge_{c \in \phi_p} c \setminus \{ p \} \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \neg(c \setminus \{ p \}) \right]_1^{w(p)} \\
                                                               &= \left[ \bigvee_{c \in \phi_p} \neg \bigvee_{l \in c \setminus \{ p \}} l \right]_1^{w(p)} = \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l \right]_1^{w(p)}.
  \end{align*}
   By \cref{cond:disjoint}, $\bigwedge_{l \in c \setminus \{ p \}} \neg l$ can
   be true for at most one $c \in \phi_p$, and so
   \[
     \left[ \bigvee_{c \in \phi_p} \bigwedge_{l \in c \setminus \{ p \}} \neg l
     \right]_1^{w(p)} = \prod_{c \in \phi_p} \left[ \bigwedge_{l \in c \setminus
         \{ p \}} \neg l \right]_1^{w(p)}
   \]
   which is exactly what we needed to show. This ends the proof that the
   first loop of \cref{alg:transformation} preserves the answer under both
   \cref{condition:d02} and \cref{condition:bklm16}. Finally, the loop on
   \crefrange{line:foreach2start}{line:foreach2end} of \cref{alg:transformation}
   replaces $[v]_1^p[\neg v]_1^q$ with $[v]_q^p$ (for some $v \in X_I$ and $p, q
   \in \mathbb{R}$), but, of course,
   \[
     [v]_1^p[\neg v]_1^q = [v]_1^p[v]_q^1 = [v]_q^p,
   \]
   i.e., the answer is unchanged.
\end{proof}

@@

\paragraph{Notes.}
\begin{itemize}
\item The experiments were run on a computing cluster with Intel Xeon E5-2630,
  Intel Xeon E7-4820, and Intel Xeon Gold 6138 processors with a
  \SI{1000}{\second} timeout separately on both encoding and inference, and a
  \SI{32}{\gibi\byte} memory limit.\footnote{Each instance was run on the same
    processor across all algorithms and encodings.}
\item Along with
  \textsf{DPMC}\footnote{\url{https://github.com/vardigroup/DPMC}}, in our
  experiments we also include WMC algorithms used in the papers that introduce
  each encoding: \textsf{Ace}
  for \texttt{cd05}, \texttt{cd06}, and \texttt{d02};
  \textsf{Cachet}\footnote{\url{https://cs.rochester.edu/u/kautz/Cachet/}}
  \cite{DBLP:conf/sat/SangBBKP04} for \texttt{sbk05}; and
  \textsf{c2d}\footnote{\url{http://reasoning.cs.ucla.edu/c2d/}}
  \cite{DBLP:conf/ecai/Darwiche04} with
  \textsf{query-dnnf}\footnote{\url{http://www.cril.univ-artois.fr/kc/d-DNNF-reasoner.html}}
  for \texttt{bklm16}.
\item \textsf{DPMC} was run in \emph{greedy mode}. This mode (which was not part
  of the original paper \cite{DBLP:conf/cp/DudekPV20}) optimises the order in
  which pseudo-Boolean functions are multiplied by prioritising functions with
  small representations.
\item Indeed, the numbers of variables in \texttt{bklm16++} and \texttt{d02++}
  encodings are equal whenever all variables in a Bayesian network have two
  values (which applies to most of the benchmark instances). Nonetheless,
  \texttt{bklm16++} performs much faster than \texttt{d02++}.
\item \textsf{DPMC} is run with tree decomposition based planning and algebraic
  decision diagram based execution---the best-performing combination in the
  original set of experiments \cite{DBLP:conf/cp/DudekPV20}. We use a single
  iteration of \textsf{htd} \cite{DBLP:conf/cpaior/AbseherMW17} to generate
  approximately optimal tree decompositions---we found that this configuration
  is efficient enough to handle huge instances, and yet the width of the
  returned decomposition is unlikely to differ from optimal by more than one or
  two.
\item We also record numbers of variables in each encoding before and after our
  additional transformation.
\item Experimental research questions:
  \begin{itemize}
  \item Can parameter variable elimination improve inference speed?
  \item How does DPMC combined with encodings without (and with) parameter
    variables compare with other WMC algorithms and other encodings?
  \item Which instances is our approach particularly successful on (compared to
    other algorithms and encodings and to the same encoding before our
    transformation)?
  \item What proportion of variables is typically removed?
  \item Do some encodings benefit from this transformation more than others?
  \end{itemize}
\item TODO: How many times am I faster than old cd06 and new bklm16? Perhaps add
  other numbers as well.
\item To perform Bayesian network inference with \textsf{DPMC} (or with any
  other WMC algorithm not based on compilation such as \textsf{Cachet}), one
  needs to select a probability to compute
  \cite{DBLP:conf/cp/DudekPV20,DBLP:conf/sat/SangBBKP04}. If a network comes
  with an evidence file, we compute the probability of this evidence. Otherwise,
  let $X$ be the variable last mentioned in the Bayesian network file. If
  \textsf{true} is one of the values of $X$, then we compute $\Pr(X =
  \textsf{true})$, otherwise we choose the first-mentioned value of $X$.
\end{itemize}
% TODO: describe the results.
